Binocular integration of chromatic and luminance signals

### 2024-04-18 manuscript.qmd
Fixed python
Update to threshold criteria

if (processdata>0){
  # install_python(version="3.6")
  # py_install('https://github.com/wichmann-lab/python-psignifit/zipball/master')

  use_miniconda('r-reticulate')

@@

@@

modelString = "
  data {
    int<lower=1> Ndatasets ;
    int<lower=1> Ntotal ;
    int ncorrect[Ntotal] ;
    int ntrials[Ntotal] ;
    real PedestalContrast[Ntotal] ;
    real TargetContrast[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;  
  }
  parameters {
    vector<lower=1, upper=3>[Ndatasets] p;
    vector<lower=2, upper=16>[Ndatasets] q;
    vector<lower=0.5, upper=2>[Ndatasets] m;
    vector<lower=0, upper=2>[Ndatasets] S;  
    vector<lower=0, upper=1>[Ndatasets] Z;
    vector<lower=0, upper=2>[Ndatasets] w;
    vector<lower=0, upper=2>[Ndatasets] k;

@@

    dprime = (targresp-nullresp)/k[s[i]] ;
    if (is_nan(dprime)){
       dprime = 0;
    }
    if (is_inf(dprime)){
       dprime = 10;
    }
    if (dprime > 10){
       dprime = 10;
    }    
    ncorrect[i] ~ binomial(ntrials[i], Phi(dprime/sqrt(2))) ;
    }
  }  
" # close quote for modelString

@@

    targresp = pow((Lt + Rt), ",allstanparams[1,1],") / (ZAc[s[i]] + pow((Lt + Rt), ",allstanparams[1,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[1,1],") / (ZAc[s[i]] + pow((Ln + Rn), ",allstanparams[1,2],")) ;     
    dprime = (targresp-nullresp)/",allstanparams[1,7]," ;

@@

    targresp = pow((Lt + Rt), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Lt + Rt), ",allstanparams[2,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Ln + Rn), ",allstanparams[2,2],")) ;     
    dprime = (targresp-nullresp)/",allstanparams[2,7]," ;

@@

    targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,2],")) ;     
    dprime = (targresp-nullresp)/",allstanparams[3,7]," ;

@@

      targresp = pow((Lt + Rt), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Lt + Rt), ",allstanparams[2,2],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Ln + Rn), ",allstanparams[2,2],")) ;     
      dprime = (targresp-nullresp)/",allstanparams[2,7]," ;
    }

@@

      targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,2],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,2],")) ;     
      dprime = (targresp-nullresp)/",allstanparams[3,7]," ;
    }

@@

      targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,4],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,4],")) ;     
      dprime = (targresp-nullresp)/",allstanparams[3,7]," ;
    }

@@

  tr <- getmodelresp3inputs(p,contrastsC,0,0,0,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[1,] - nr[1,])/p[1,7]
  pmfout[1,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,0,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[2,] - nr[2,])/p[2,7]
  pmfout[2,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,0,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[3,] - nr[3,])/p[3,7]
  pmfout[3,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,16,0,0,0,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[1,] - nr[1,])/p[1,7]
  pmfout[4,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,16,contrastsC,0,0,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[2,] - nr[2,])/p[2,7]
  pmfout[5,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,16,0,0,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[3,] - nr[3,])/p[3,7]
  pmfout[6,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,0,0,16,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[1,] - nr[1,])/p[1,7]
  pmfout[7,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,16,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[2,] - nr[2,])/p[2,7]
  pmfout[8,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,16,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[3,] - nr[3,])/p[3,7]
  pmfout[9,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,0,0,0,0,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[1,] - nr[1,])/p[1,7]
  pmfout[10,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,0,0,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[2,] - nr[2,])/p[2,7]
  pmfout[11,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,0,contrastsC,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[3,] - nr[3,])/p[3,7]
  pmfout[12,] <- 100*pnorm(dprime/sqrt(2))

@@

\begin{equation}
Stage1_L = \frac{C_L^m}{S + C_L + \omega C_R},
\end{equation}

\begin{equation}
Stage1_R = \frac{C_R^m}{S + C_R + \omega C_L},
\end{equation}

\begin{equation}
binsum = Stage1_L + Stage1_R,
\end{equation}

\begin{equation}
Stage2 = \frac{binsum^p}{Z + binsum^q},
\end{equation}

where $C_L$ and $C_R$ are the (normalized) contrasts displayed to the left and right eyes, and $m$, $S$, $\omega$, $p$, $q$ and $Z$ are free parameters in the model. A further free parameter, $k$, represents additive internal noise, and is used to convert the model outputs to either d-prime or threshold values (note that in the original model specification [@Meese2006] this parameter was called $\sigma$, but we use the $k$ symbol here to avoid confusion with the standard deviation of the cumulative Gaussian used when fitting the psychometric functions to estimate thresholds). Thresholds are defined by iteratively adjusting the target contrast until the following equality is satisfied:

\begin{equation}
Stage2_{target+pedestal} - Stage2_{pedestal} = k,
\end{equation}

and d-prime for a single target level is defined as:

\begin{equation}
d' = \frac{Stage2_{target+pedestal} - Stage2_{pedestal}}{k}.
\end{equation}

@@

\begin{equation}
ACStage1_L = \frac{AC_L^m}{S + AC_L + \omega_A AC_R + \omega_R RG_R + \omega_B BY_R},
\end{equation}

@@

For consistency with previous work, we initially performed least-squares fits of the two-stage model [@Meese2006] for each of the chromaticity experiments, and the flickering disc experiment (7 free parameters per fit). The best model fits are shown by the curves in @fig-dipperfig and @fig-discdata, and provide an excellent description of the data, with RMS errors between `r round(min(allrms),digits=2)` and `r round(max(allrms),digits=2)`dB. Best fitting parameters are shown in the 'Simplex fits' section of @tbl-parametertable. We note that in previous work, the weight of interocular suppression ($\omega$ in the model) was implicitly fixed at 1. Here we allowed it to vary, but it still received a value very close to 1 in all of our grating conditions, and slightly below 1 (`r round(allp[4,6],digits=2)`) for our flickering discs. The exponent values (*p* and *q*) for the blue/yellow gratings are rather different from those of the other conditions. In previous work [@Meese2006] the second stage gain control nonlinearity (Equation 4) does typically have quite substantial exponent values, which balance the relatively mild nonlinearity at the first stage, and produce a compressive transducer that results in contrast masking (the handle of the dipper). The high value of *p* = `r round(allp[3,1], digits=2)` is therefore likely to be compensating for the low value of *m* = `r round(allp[3,3], digits=2)` at stage one, but may well represent the combination of several successive stages of nonlinearity, and perhaps also other phenomena such as uncertainty [@Pelli1985].

@@

Estimates of the binocular summation ratio at threshold fell within the range $\sqrt{2}$ to 2 for all four stimuli tested here. Consistent with previous reports [@Simmons2005], summation was slightly higher for chromatic versus achromatic stimuli, but fell short of perfect linear summation (a ratio of 2). In our model, summation at threshold is determined by the exponent at the first gain control stage (the *m* parameter in Equations 1 & 2). The summation ratio can be approximated by $2^{1/m}$ [see Figure 9 of @Baker2012], such that an exponent of $m=1$ produces a ratio of 2, and an exponent of $m = 2$ produces a ratio of $\sqrt{2}$. Both the simplex and Bayesian model fits (see @tbl-parametertable) generated parameter estimates in the range $1 < m < 1.4$, consistent with the high levels of summation observed empirically.

@@

The weight of interocular suppression is represented by the model parameter, $\omega$ (Equations 1 & 2). Posterior distributions of this parameter overlapped 1 for all of our dipper function data sets (see top right panel in [@fig-bayesianmodel]), indicating strong interocular suppression regardless of post-retinal pathway. The flickering disc stimuli produced a slightly lower suppression estimate, with $\omega \sim 0.9$. This may indicate slightly weaker suppression between the eyes for temporal modulations, but it is much less extreme than our recent estimates using EEG and matching paradigms [@Segala2023]. One difference between these studies is the luminance of the background, which was set to black for the experiments of @Segala2023, but here was set at the mean luminance. Other studies using static luminance increments have also reported differences in the character of binocular combination that are attributable to background luminance [@Baker2012], so this may explain the differences between studies. We also note that the weaker interocular suppression for flickering disc stimuli appears to be largely due to one participant (see [@fig-individualdiscs]c), so individual differences might also play a role. Future work should manipulate the background luminance systematically to better understand how this modulates interocular suppression.

@@

Recent evidence indicates that the physiological substrate of interocular suppression may be neurons in layer 4 of primary visual cortex [@Dougherty2019]. Most cells in this layer are monocularly excitable, in that their responses increase only by stimulation of their preferred eye. However, simultaneous stimulation of the non-preferred eye can modulate the response, usually in an inhibitory fashion, exactly as proposed at stage 1 of the two-stage model (Equations 1 & 2). In terms of perception, one consequence of this early suppression is to achieve 'ocularity invariance', whereby the perceived contrast of a stimulus viewed by one eye is equivalent to that of the same stimulus viewed by both eyes [@Baker2007]. Similar processes of response invariance have also been reported using fMRI [@Moradi2009] and steady-state EEG [@Baker2017].

>>>>

if (processdata>0){
  install_python(version="3.6")
  py_install('https://github.com/wichmann-lab/python-psignifit/zipball/master')

  # use_miniconda('r-reticulate')

@@

# calculate d-prime at 75% correct threshold
thdcriteria <- qnorm(0.75)*sqrt(2)

@@

modelString = paste0("
  data {
    int<lower=1> Ndatasets ;
    int<lower=1> Ntotal ;
    int ncorrect[Ntotal] ;
    int ntrials[Ntotal] ;
    real PedestalContrast[Ntotal] ;
    real TargetContrast[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;  
  }
  parameters {
    vector<lower=1, upper=3>[Ndatasets] p;
    vector<lower=2, upper=16>[Ndatasets] q;
    vector<lower=0.5, upper=2>[Ndatasets] m;
    vector<lower=0, upper=2>[Ndatasets] S;  
    vector<lower=0, upper=1>[Ndatasets] Z;
    vector<lower=0, upper=2>[Ndatasets] w;
    vector<lower=0, upper=2>[Ndatasets] k;

@@

    dprime = (targresp-nullresp)/(k[s[i]]/",thdcriteria,") ;
    if (is_nan(dprime)){
       dprime = 0;
    }
    if (is_inf(dprime)){
       dprime = 10;
    }
    if (dprime > 10){
       dprime = 10;
    }    
    ncorrect[i] ~ binomial(ntrials[i], Phi(dprime/sqrt(2))) ;
    }
  }  
") # close quote for modelString

@@

    targresp = pow((Lt + Rt), ",allstanparams[1,1],") / (ZAc[s[i]] + pow((Lt + Rt), ",allstanparams[1,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[1,1],") / (ZAc[s[i]] + pow((Ln + Rn), ",allstanparams[1,2],")) ;     
    dprime = (targresp-nullresp)/(",allstanparams[1,7],"/",thdcriteria,") ;

@@

    targresp = pow((Lt + Rt), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Lt + Rt), ",allstanparams[2,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Ln + Rn), ",allstanparams[2,2],")) ;     
    dprime = (targresp-nullresp)/(",allstanparams[2,7],"/",thdcriteria,") ;

@@

    targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,2],")) ;
    nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,2],")) ;     
    dprime = (targresp-nullresp)/(",allstanparams[3,7],"/",thdcriteria,") ;

@@

      targresp = pow((Lt + Rt), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Lt + Rt), ",allstanparams[2,2],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[2,1],") / (ZRg[s[i]] + pow((Ln + Rn), ",allstanparams[2,2],")) ;     
      dprime = (targresp-nullresp)/(",allstanparams[2,7],"/",thdcriteria,") ;
    }

@@

      targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,2],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,2],")) ;     
      dprime = (targresp-nullresp)/(",allstanparams[3,7],"/",thdcriteria,") ;
    }

@@

      targresp = pow((Lt + Rt), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Lt + Rt), ",allstanparams[3,4],")) ;
      nullresp = pow((Ln + Rn), ",allstanparams[3,1],") / (ZBy[s[i]] + pow((Ln + Rn), ",allstanparams[3,4],")) ;     
      dprime = (targresp-nullresp)/(",allstanparams[3,7],"/",thdcriteria,") ;
    }

@@

  tr <- getmodelresp3inputs(p,contrastsC,0,0,0,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[1,] - nr[1,])/(p[1,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[1,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,0,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[2,] - nr[2,])/(p[2,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[2,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,0,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,0)
  dprime <- (tr[3,] - nr[3,])/(p[3,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[3,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,16,0,0,0,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[1,] - nr[1,])/(p[1,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[4,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,16,contrastsC,0,0,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[2,] - nr[2,])/(p[2,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[5,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,16,0,0,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,16,0,0,0,0)
  dprime <- (tr[3,] - nr[3,])/(p[3,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[6,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,0,0,16,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[1,] - nr[1,])/(p[1,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[7,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,16,0,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[2,] - nr[2,])/(p[2,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[8,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,16,contrastsC,0)
  nr <- getmodelresp3inputs(p,0,0,0,16,0,0)
  dprime <- (tr[3,] - nr[3,])/(p[3,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[9,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,contrastsC,0,0,0,0,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[1,] - nr[1,])/(p[1,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[10,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,contrastsC,0,0,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[2,] - nr[2,])/(p[2,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[11,] <- 100*pnorm(dprime/sqrt(2))

  tr <- getmodelresp3inputs(p,0,0,0,0,contrastsC,16)
  nr <- getmodelresp3inputs(p,0,0,0,0,0,16)
  dprime <- (tr[3,] - nr[3,])/(p[3,7]/(qnorm(0.75)*sqrt(2)))
  pmfout[12,] <- 100*pnorm(dprime/sqrt(2))

@@

$$\mathrm{Stage1}_L = \frac{C_L^m}{S + C_L + \omega C_R},$${#eq-s1l}

$$\mathrm{Stage1}_R = \frac{C_R^m}{S + C_R + \omega C_L},$${#eq-s1r}

$$\mathrm{binsum} = \mathrm{Stage1}_L + \mathrm{Stage1}_R,$${#eq-binsum}

$$\mathrm{Stage2} = \frac{\mathrm{binsum}^p}{Z + \mathrm{binsum}^q},$${#eq-s2}

where $C_L$ and $C_R$ are the (normalized) contrasts displayed to the left and right eyes, and $m$, $S$, $\omega$, $p$, $q$ and $Z$ are free parameters in the model. A further free parameter, $k$, is used to convert the model outputs to either d-prime or threshold values (note that in the original model specification [@Meese2006] this parameter was called $\sigma$, but we use the $k$ symbol here to avoid confusion with the standard deviation of the cumulative Gaussian used when fitting the psychometric functions to estimate thresholds). Thresholds are defined by iteratively adjusting the target contrast until the following equality is satisfied:

$$\mathrm{Stage2}_{\mathrm{target+pedestal}} - \mathrm{Stage2}_{\mathrm{pedestal}} = k,$${#eq-thd}

and $d'$ (d-prime) for a single target level is defined as:

$$d' = \frac{\mathrm{Stage2}_{\mathrm{target+pedestal}} - \mathrm{Stage2}_{\mathrm{pedestal}}}{k/\tau},$${#eq-dprime}

where the parameter $\tau$ reflects the value of $d'$ at detection threshold. Defining threshold as the 75% correct point on the psychometric function (as here) yields a value of $\tau = \Phi^{-1}(0.75)\sqrt{2}$ = `r round(qnorm(0.75)*sqrt(2),digits=3)` (where $\Phi^{-1}$ is the inverse cumulative normal density function). The combined denominator term ($k/\tau$) therefore represents internal additive noise in the model.

@@

$$^{AC}\mathrm{Stage1}_L = \frac{AC_L^m}{S + AC_L + \omega_A AC_R + \omega_R RG_R + \omega_B BY_R},$${#eq-s1multi}

@@

For consistency with previous work, we initially performed least-squares fits of the two-stage model [@Meese2006] for each of the chromaticity experiments, and the flickering disc experiment (7 free parameters per fit). The best model fits are shown by the curves in @fig-dipperfig and @fig-discdata, and provide an excellent description of the data, with RMS errors between `r round(min(allrms),digits=2)` and `r round(max(allrms),digits=2)`dB. Best fitting parameters are shown in the 'Simplex fits' section of @tbl-parametertable. We note that in previous work, the weight of interocular suppression ($\omega$ in the model) was implicitly fixed at 1. Here we allowed it to vary, but it still received a value very close to 1 in all of our grating conditions, and slightly below 1 (`r round(allp[4,6],digits=2)`) for our flickering discs. The exponent values (*p* and *q*) for the blue/yellow gratings are rather different from those of the other conditions. In previous work [@Meese2006] the second stage gain control nonlinearity (@eq-s2) does typically have quite substantial exponent values, which balance the relatively mild nonlinearity at the first stage, and produce a compressive transducer that results in contrast masking (the handle of the dipper). The high value of *p* = `r round(allp[3,1], digits=2)` is therefore likely to be compensating for the low value of *m* = `r round(allp[3,3], digits=2)` at stage one, but may well represent the combination of several successive stages of nonlinearity, and perhaps also other phenomena such as uncertainty [@Pelli1985].

@@

Estimates of the binocular summation ratio at threshold fell within the range $\sqrt{2}$ to 2 for all four stimuli tested here. Consistent with previous reports [@Simmons2005], summation was slightly higher for chromatic versus achromatic stimuli, but fell short of perfect linear summation (a ratio of 2). In our model, summation at threshold is determined by the exponent at the first gain control stage (the *m* parameter in @eq-s1l and @eq-s1r). The summation ratio can be approximated by $2^{1/m}$ [see Figure 9 of @Baker2012], such that an exponent of $m=1$ produces a ratio of 2, and an exponent of $m = 2$ produces a ratio of $\sqrt{2}$. Both the simplex and Bayesian model fits (see @tbl-parametertable) generated parameter estimates in the range $1 < m < 1.4$, consistent with the high levels of summation observed empirically.

@@

The weight of interocular suppression is represented by the model parameter, $\omega$ (@eq-s1l and @eq-s1r). Posterior distributions of this parameter overlapped 1 for all of our dipper function data sets (see top right panel in [@fig-bayesianmodel]), indicating strong interocular suppression regardless of post-retinal pathway. The flickering disc stimuli produced a slightly lower suppression estimate, with $\omega \sim 0.9$. This may indicate slightly weaker suppression between the eyes for temporal modulations, but it is much less extreme than our recent estimates using EEG and matching paradigms [@Segala2023]. One difference between these studies is the luminance of the background, which was set to black for the experiments of @Segala2023, but here was set at the mean luminance. Other studies using static luminance increments have also reported differences in the character of binocular combination that are attributable to background luminance [@Baker2012], so this may explain the differences between studies. We also note that the weaker interocular suppression for flickering disc stimuli appears to be largely due to one participant (see [@fig-individualdiscs]c), so individual differences might also play a role. Future work should manipulate the background luminance systematically to better understand how this modulates interocular suppression.

@@

Recent evidence indicates that the physiological substrate of interocular suppression may be neurons in layer 4 of primary visual cortex [@Dougherty2019]. Most cells in this layer are monocularly excitable, in that their responses increase only by stimulation of their preferred eye. However, simultaneous stimulation of the non-preferred eye can modulate the response, usually in an inhibitory fashion, exactly as proposed at stage 1 of the two-stage model (@eq-s1l and @eq-s1r). In terms of perception, one consequence of this early suppression is to achieve 'ocularity invariance', whereby the perceived contrast of a stimulus viewed by one eye is equivalent to that of the same stimulus viewed by both eyes [@Baker2007]. Similar processes of response invariance have also been reported using fMRI [@Moradi2009] and steady-state EEG [@Baker2017].
